---
title: "methods"
author: "Yuqi Miao ym2771"
date: "3/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = F)
```

# Methods

## Data

The mean statistics for every feature are selected as predictors in the model. The sample size for the data is 569. The response variable is  "diagnosis", which is transformed to a binary variable (Benign = 0, Malignant = 1),and the 11 predictors are standardised. 

```{r}
cancer = read_csv("breast-cancer.csv") %>% 
  mutate(diagnosis = as.numeric(factor(diagnosis, levels = c("B","M"), labels = c(0,1)))-1) %>% 
  mutate(radius_mean = scale(radius_mean))
predictor_scale = as.tibble(scale(cancer[3:12]))
cancer_scale = cbind(rep(1,569),predictor_scale, cancer$diagnosis)
names(cancer_scale) = c("ones", names(cancer)[3:12], "response")
# head(cancer_scale)
## standardize data!!
```

## Model define

For $i_{th}$ observation, the response variable $Y_i$ follows binary distribution:

$$Y_i\sim Bin(\pi_i)$$

the log-liklihood function:

$$l(\mathbf Y,\boldsymbol\pi)=\sum_{i=1}^n l(y_i,\pi_i)=\sum_{i=1}^n(y_ilog\frac{\pi_i}{1-\pi_i}+log(1-\pi_i))\\=\sum_{i=1}^n(y_ilog\frac{\pi_i}{1-\pi_i}+log(1-\pi_i))$$

Where $\pi_i$ denotes the probablity of the $i_{th}$ observation to be maglinant.

To build relationship between the response and predictors, the GLM is defined as:

$$\log(\frac{\pi_i}{1-\pi_i}) = \mathbf x_i \boldsymbol \beta= \theta_i$$



## Full model

Firstly, Newton-Rapson method is used to fit the full model. To find the maximum likelihood estimation of coefficients, iteration process is set as follows:

$$\theta_{i+1}  = \theta_{i} -(\nabla^2l(\theta_{i}|\boldsymbol X))^{-1}\nabla l(\theta_{i}|\boldsymbol X) $$

$$\nabla l(\theta|\boldsymbol X)=(\frac{\partial\boldsymbol \theta}{\partial \boldsymbol\beta})^T\times \frac{\partial l(y,\beta)}{\partial \boldsymbol \theta}\\=\mathbf X^T(\mathbf Y-\frac{e^\boldsymbol \theta}{1+e^\boldsymbol \theta})$$

$$\nabla^2 l(\theta|\boldsymbol X)=-\mathbf X^Tdiag(\pi_i(1-\pi_i))\mathbf X$$






















